###331-70 Final Project

####Ajay Patel, Ethan Moore, and Jack Langston

#####Intro

For our project, we decided to create a Shiny app to visualize the results of the 2018 Major League Baseball Postseason. Our app allows users to interact with the data we downloaded from BaseballSavant.com. The dataset includes dozens of variables for each of the approximately 10,000 pitches from the postseason, including pitch type, whether or not the ball was hit, what event occurred as a result of the hit, etc.

Our graph is a display of the pitch zone as seen from the catcher's perspective. The black box in the middle of our display represents the strike zone. Users can select a player, an opposing team, and a pitch type and our app will show the graph of each pitch of the selected pitch type seen by the selected player against the selected team as measured by an all-encompassing offensive metric called ["wOBA"](https://www.fangraphs.com/library/offense/woba/). The "boxes" shown reflect the  horizontal and vertical locations of the pitches seen, and their color shows how well the hitter did on pitches in that region (according to the legend on the righthand side). 

The work on this project was collaborative with each of the group members contributing equal amounts in a variety of ways. Without all three members, this project would have looked much different and would have been much worse.

###Example 1

Let's look into how Yankee slugger Aaron Judge did against changeups thrown by the rival Red Sox:

```{r, echo = FALSE}
load("BBWorkspace.RData")
suppressWarnings(library(tidyverse))
suppressWarnings(library(RColorBrewer))

e <- is.na(our_data$events)
    our_data <- subset(our_data, e == F)
    
    
    our_data <- subset(our_data,
                       player_name == "Aaron Judge")
    
    our_data <- subset(our_data,
                         opp_team == "BOS")

    our_data <- subset(our_data,
                         pitch_type == "Changeup")
      
    
    
        
    for (i in 1:length(our_data$pitch_type)) {
      if (our_data$inning_topbot[i] == "Bot") {
        our_data$team[i] <- our_data$home_team[i]
        our_data$opp_team[i] <- our_data$away_team[i]
      } else {
        our_data$team[i] <- our_data$away_team[i]
        our_data$opp_team[i] <- our_data$home_team[i]
      }
    }
    
    our_data$pitch_type[(our_data$pitch_type) == "CH"] <- "Changeup"
    our_data$pitch_type[(our_data$pitch_type) == "CU"] <- "Curveball"
    our_data$pitch_type[(our_data$pitch_type) == "FC"] <- "Cutter"
    our_data$pitch_type[(our_data$pitch_type) == "FF"] <- "4-Seam Fastball"
    our_data$pitch_type[(our_data$pitch_type) == "FS"] <- "Splitter"
    our_data$pitch_type[(our_data$pitch_type) == "FT"] <- "Two-Seam Fastball"
    our_data$pitch_type[(our_data$pitch_type) == "KC"] <- "Knuckle Curveball"
    our_data$pitch_type[(our_data$pitch_type) == "SI"] <- "Sinker"
    our_data$pitch_type[(our_data$pitch_type) == "SL"] <- "Slider"
    
    our_data$player_name <- as.factor(our_data$player_name)
    our_data$opp_team <- as.factor(our_data$opp_team)
    our_data$team <- as.factor(our_data$team)
    our_data$pitch_type <- as.factor(our_data$pitch_type)
    
    x_zoning <- function(x) {
      if (is.na(x)){
        return (0)
      } else if (x <= -0.708){
        return(1)
      } else if (x > -0.708 & x <= -0.2357) {
        return(2)
      } else if (x > -0.2357 & x <= 0.2357) {
        return(3)
      } else if (x > .2357 & x <= .708) {
        return(4) 
      } else if (x > .708) {
        return(5)
      } else {
        return(0)
      }
    }
    
    z_zoning <- function(z) {
      if (is.na(z)) {
        return (0)
      } else if (z <= 1.5) {
        return(1)
      } else if (z > 1.5 & z <= 2.167) {
        return(2)
      } else if (z > 2.167 & z <= 2.8337) {
        return(3)
      } else if (z > 2.8337 & z <= 3.5) {
        return(4) 
      } else if (z > 3.5) {
        return(5)
      } else {
        return(0)
      }
    }
    
    for (i in 1:length(our_data$pitch_type)){
      our_data$x_zone[i] <- x_zoning(our_data$plate_x[i])
      our_data$z_zone[i] <- z_zoning(our_data$plate_z[i])
    }
    
    for (i in 1:length(our_data$pitch_type)){
      if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 11
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 12
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 13
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 14
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 15
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 21
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 22
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 23
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 24
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 25
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 31
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 32
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 33
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 34
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 35
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 41
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 42
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 43
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 44
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 45
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 51
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 52
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 53
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 54
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 55
      } else {
        our_data$doub_zone[i] = 99
      }
    }
    
    bin_woba_value <- our_data %>%
      group_by(doub_zone) %>%
      summarise(bin_woba_value = mean(woba_value))
    wobas <- bin_woba_value

    for (i in 1:length(our_data$doub_zone)) { 
      for (j in 1:length(bin_woba_value$doub_zone)) {
        if (our_data$doub_zone[i] == bin_woba_value$doub_zone[j]) {
          our_data$bin_woba[i] <- bin_woba_value$bin_woba_value[j]
        }
      }
    }
    mycolors <- brewer.pal(3, "Reds")
    ggplot(data = our_data,
           aes(x = x_zone,
               y = z_zone,
               fill = bin_woba)) +
      geom_tile() +
      #scale_fill_gradient(low = 'white', high = 'red') +
      scale_fill_gradient(low = mycolors[1], high = mycolors[3]) +
      xlim(0.5,5.5) +
      ylim(0.5,5.5) 
    
  
```

This graph shows that Judge saw changeups against the Red Sox in two regions, both high and inside, and had the same results on all of them. A wOBA of 0.9 is very good considering the average wOBA for a player in the postseason this year was 0.296.

###Example 2

In the National League, the Los Angeles Dodgers made it all the way to the World Series, in part due to their mid-season acquisition of third-baseman Manny Machado. 

```{r, echo = F}
load("BBWorkspace.RData")

e <- is.na(our_data$events)
    our_data <- subset(our_data, e == F)
    
    
    our_data <- subset(our_data,
                       player_name == "Manny Machado")
    
    our_data <- subset(our_data,
                         opp_team == "BOS")

    our_data <- subset(our_data,
                         pitch_type == "4-Seam Fastball")
      
    
    
        
    for (i in 1:length(our_data$pitch_type)) {
      if (our_data$inning_topbot[i] == "Bot") {
        our_data$team[i] <- our_data$home_team[i]
        our_data$opp_team[i] <- our_data$away_team[i]
      } else {
        our_data$team[i] <- our_data$away_team[i]
        our_data$opp_team[i] <- our_data$home_team[i]
      }
    }
    
    our_data$pitch_type[(our_data$pitch_type) == "CH"] <- "Changeup"
    our_data$pitch_type[(our_data$pitch_type) == "CU"] <- "Curveball"
    our_data$pitch_type[(our_data$pitch_type) == "FC"] <- "Cutter"
    our_data$pitch_type[(our_data$pitch_type) == "FF"] <- "4-Seam Fastball"
    our_data$pitch_type[(our_data$pitch_type) == "FS"] <- "Splitter"
    our_data$pitch_type[(our_data$pitch_type) == "FT"] <- "Two-Seam Fastball"
    our_data$pitch_type[(our_data$pitch_type) == "KC"] <- "Knuckle Curveball"
    our_data$pitch_type[(our_data$pitch_type) == "SI"] <- "Sinker"
    our_data$pitch_type[(our_data$pitch_type) == "SL"] <- "Slider"
    
    our_data$player_name <- as.factor(our_data$player_name)
    our_data$opp_team <- as.factor(our_data$opp_team)
    our_data$team <- as.factor(our_data$team)
    our_data$pitch_type <- as.factor(our_data$pitch_type)
    
    x_zoning <- function(x) {
      if (is.na(x)){
        return (0)
      } else if (x <= -0.708){
        return(1)
      } else if (x > -0.708 & x <= -0.2357) {
        return(2)
      } else if (x > -0.2357 & x <= 0.2357) {
        return(3)
      } else if (x > .2357 & x <= .708) {
        return(4) 
      } else if (x > .708) {
        return(5)
      } else {
        return(0)
      }
    }
    
    z_zoning <- function(z) {
      if (is.na(z)) {
        return (0)
      } else if (z <= 1.5) {
        return(1)
      } else if (z > 1.5 & z <= 2.167) {
        return(2)
      } else if (z > 2.167 & z <= 2.8337) {
        return(3)
      } else if (z > 2.8337 & z <= 3.5) {
        return(4) 
      } else if (z > 3.5) {
        return(5)
      } else {
        return(0)
      }
    }
    
    for (i in 1:length(our_data$pitch_type)){
      our_data$x_zone[i] <- x_zoning(our_data$plate_x[i])
      our_data$z_zone[i] <- z_zoning(our_data$plate_z[i])
    }
    
    for (i in 1:length(our_data$pitch_type)){
      if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 11
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 12
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 13
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 14
      } else if (our_data$x_zone[i] == 1 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 15
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 21
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 22
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 23
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 24
      } else if (our_data$x_zone[i] == 2 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 25
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 31
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 32
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 33
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 34
      } else if (our_data$x_zone[i] == 3 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 35
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 41
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 42
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 43
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 44
      } else if (our_data$x_zone[i] == 4 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 45
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 1){
        our_data$doub_zone[i] = 51
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 2){
        our_data$doub_zone[i] = 52
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 3){
        our_data$doub_zone[i] = 53
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 4){
        our_data$doub_zone[i] = 54
      } else if (our_data$x_zone[i] == 5 & our_data$z_zone[i] == 5){
        our_data$doub_zone[i] = 55
      } else {
        our_data$doub_zone[i] = 99
      }
    }
    
    bin_woba_value <- our_data %>%
      group_by(doub_zone) %>%
      summarise(bin_woba_value = mean(woba_value))
    wobas <- bin_woba_value

    for (i in 1:length(our_data$doub_zone)) { 
      for (j in 1:length(bin_woba_value$doub_zone)) {
        if (our_data$doub_zone[i] == bin_woba_value$doub_zone[j]) {
          our_data$bin_woba[i] <- bin_woba_value$bin_woba_value[j]
        }
      }
    }
    mycolors <- brewer.pal(3, "Reds")
    ggplot(data = our_data,
           aes(x = x_zone,
               y = z_zone,
               fill = bin_woba)) +
      geom_tile() +
      #scale_fill_gradient(low = 'white', high = 'red') +
      scale_fill_gradient(low = mycolors[1], high = mycolors[3]) +
      xlim(0.5,5.5) +
      ylim(0.5,5.5) 
```

In World Series games (against the Red Sox), this graph shows how Machado did against 4-seam fastballs. He saw fastballs in seven different zones, getting a hit in only one zone, unsurprisingly in the most central zones where fastballs are often the easiest to hit.

If users select a combination of inputs that had 0 occurences, the graph will disappear and text will appear in the main panel prompting the user to select different inputs.

###Statistical Elements

The main aspect of our app shows the wOBA for each pitch and similar pitches in similar strike zone locations. Aside from this visualization of said descriptive statistics, we ran a one-sample t-test comparing the average wOBA for the postseason and the average wOBA for the entire 2018 Regular Season, displying the results below the main graph. Overall, we discovered there is a lot of variance in how well a player does in any particular series and which pitch he saw. In one series, a given player may do well against fastballs, and then in the next, it's like same player has never played Baseball before.  








